---
title: 'Data Visualization: Tables'
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures",
  out.width = "100%"
)
```

```{r LIBRARIES, message=FALSE, warning=FALSE}
library(data.table)
library(dplyr)

library(ggplot2)
library(scales)
library(formattable)
library(DT)
```

# T A B L E

One thing to keep in mind with a table is that you want the design to
fade into the background, letting the data take center stage. Don't let
heavy borders or shading compete for attention. Instead, think of using
light borders or simply white space to set apart elements of the table.

Borders should be used to improve the legibility of your table. Think
about pushing them to the background by making them grey, or getting rid
of them altogether. The data should be what stands out, not the borders.

```{r SIMPLE TABLE}
data <- head(airquality[, c("Day", "Temp")], 10)

# Define a simple table
simple_table <- formattable(data,
                            col.names = c("Day", "Temp"),
                            align = c("c", "c"))

# Adjust the simple tables' width
as.htmlwidget(simple_table, width = 400)
```

# H E A T M A P

A heatmap is a way to visualize data in tabular format, where in place
of (or in addition to) the numbers, you leverage colored cells that
convey the relative magnitude of the numbers.

To reduce this mental processing, we can use color saturation to provide
visual cues, helping our eyes and brains more quickly target the
potential points of interest.

```{r CONDITIONALLY FORMATTED CELLS}
# Define a table with conditionally formatted cells
cond_table <- formattable(data, 
                          col.names = c("Day", "Temp"),
                          list(Temp = color_tile("white", "#014991")), # Indicate the name of the column
                          align = c("c", "c")
)

# Adjust the conditionally formatted tables' width
as.htmlwidget(cond_table, width = 400)
```

```{r FORMATTER RED / GREEN}
#Create a formatter to display the red/green and arrow up/down indicators.
sign_formatter <- formatter("span", 
                            style = x ~ style(color = ifelse(x > 60, "#168346",
                                                      ifelse(x < 60, "#C3514E", "black"))))

cond_table2 <- formattable(data, 
                          col.names = c("Day", "Temp"),
                          list(Temp = sign_formatter), # Indicate the name of the column
                          align = c("c", "c"))

# Adjust the conditionally formatted tables' width
as.htmlwidget(cond_table2, width = 400)
```

```{r}
cond_table3 <- formattable(data, 
                          col.names = c("Day", "Temp"),
                          list(Temp = color_bar("#63a2eb")), # Indicate the name of the column
                          align = c("c", "c"))

# Adjust the conditionally formatted tables' width
as.htmlwidget(cond_table3, width = 400)
```
